---
title: "ECON 213 - PSET2"
output:
  pdf_document: default
  html_notebook: default
---


### Part 1 Data Cleaning 

- loaded packages
- Read avocado CSV data 
- remove added X1 column 
- convert column titles to tiles without spaces to minimize future R issues  
- Changed PLU numbers to continue PLU so R would recognize the column title 
- Used lubridate to make date elements (Month, Day, Year) their own column 
- Created new column ("season") that grouped months together based on which season they occur 
  - For temoral analysis to understand supply and price trends over time 
- created avocados_region which filters for groupings of states based on geography 
  - for spatial analysis to see pricing differences
- For calculating the Elasticity of Supply, columns were added to the time series
  prepared dataframe for percent change in quantity over last period and percent
  change in price over last period. This was made in order to best estimate 
  how the market of avocados changed over time, specificially looking at quantity
  and price. 
  
-Calculating the elasticity of demand entailed narrowing in on a certain, demand
driven event. I filtered for the witer months of January and February where there
was unlikey to be a supply side shift due to climate. From here, I could 
calculate the response between change in quantity and change in price during 
this event. 


```{r echo= FALSE}
library(ggplot2) ## plots stuff
library(tidyverse) ## data transformation
library(lubridate) 
library(dplyr)##
library(patchwork) # To display 2 charts together
library(hrbrthemes)

```


```{r  echo= FALSE}
avocados <- read_csv("avocado.csv")

```

```{r echo = FALSE}
avocado <-
  avocados %>% 
  select(-X1) %>% 
  rename(
    date = Date,
    plu4046 = `4046`,
    plu4225 = `4225`,
    plu4770 = `4770`,
    avg_price = AveragePrice,
    total_volume = `Total Volume`,
    total_bags = `Total Bags`,
    small_bags = `Small Bags`,
    large_bags = `Large Bags`,
    xl_bags = `XLarge Bags`
    ) %>% 
  mutate(date = ymd(date))


avocados_season <- avocado %>% 
  mutate(
    year = year(date),
    month = month(date),
    day = day(date), 
    season = case_when(
      month %in% c(12, 1, 2) ~ "Winter",
      month %in% c(3, 4, 5) ~ "Spring",
      month %in% c(6, 7, 8) ~ "Summer",
      month %in% c(9, 10, 11) ~ "Fall"
    ))


avocados_region <- avocado %>% filter(region == c('Midsouth','NorthernNewEngland', "SouthCentral","Southeast","West","WestTexNewMexico","Northeast"))


conventional <- avocados_season %>% filter(type == "conventional") 
con_mean <- mean(conventional$avg_price)


organic <- avocados_season %>% filter(type == "organic") 
org_mean <- mean(organic$avg_price)


```



## Avocado Market - regional price differences 

In understanding the regional price differences, we look at the 7 aggregations 
provided in the original data and look at distribution of prices for both 
organic and conventional avocados. We see that the majority of conventional 
avocados in South Central, West, and TexNewMexico are at the lower price points.
Histograms for the same types of avocados across different regions are not 
centered at the same point. Avocado prices are priced at market rates for the 
geographic area and not at a national level. 

```{r echo = FALSE}

avocados_region  %>% 
  ggplot(aes(`avg_price`, color=type)) +
  geom_histogram(binwidth = .1, fill="white") +
  ggtitle("average price of avocados by region, subsetted by type") +
  facet_grid(rows = "region")


```

## Avocado Market - type price differences 

Organic produce tends to be sold at a higher cost and at lower volumes. The 
histogram shows the distribution of each type. Additionally counts of each type
confirm that conventional avocados sell at much higher volumes and make up a 
larger portion of the avocado market. This observation will be important when 
calculating the adjusted weighted average of price in the avocado market. 

```{r echo= FALSE}
ggplot(avocados_region, aes(x=avg_price, color=type)) +
  geom_histogram(fill="white" ,position="dodge")+
  theme(legend.position="top")+
 geom_density(alpha=.2, fill="#FF6666")+ 
  geom_vline(xintercept = con_mean, linetype="dashed", 
                color = "red", size=1)+ 
  geom_vline(xintercept = org_mean, linetype="dashed", 
                color = "blue", size=1)

```

## Avocado Market - Seasonal Price

The figure below shows the the disribution of price based on the seasons. 
Avocados are grown seasonally and as such, out-of-season produce often ends up
being priced differently than in-season produce. Produce quality also potentially 
suffers seasonally so pricing is taken into account of the poorer quality. 
Spring and winter price distributions appear to be less than the mean, meaning 
that the avocado, on average, purchased during these seasons, will be less 
expensive than summer and fall avocados. 



```{r echo=FALSE}


m1 <- avocados_season %>% 
  filter(`type` == "conventional") %>% 
  ggplot(aes(`avg_price`)) +
  geom_histogram(binwidth = .08) +
  ggtitle("Seasonal price - Conventional") +
  facet_grid(rows = "season")+
  geom_vline(aes(xintercept=median(avg_price)),
            color="navy", linetype="solid", size=1)

m2 <- avocados_season %>% 
  filter(`type` == "organic") %>% 
  ggplot(aes(`avg_price`)) +
  geom_histogram(binwidth = .08) +
  ggtitle("Seasonal price - organic") +
  facet_grid(rows = "season")+
  geom_vline(aes(xintercept=median(avg_price)),
            color="darkgreen", linetype="solid", size=1)


m1 + m2

```

 
 

### Part 3
## time series 

```{r echo= FALSE}

time_series_group <- merge(organic, conventional, by = c("date", "region")) %>% 
  filter(region != "TotalUS" )

time_series_group <- time_series_group %>%
  add_column(adj_price = (time_series_group$total_volume.x* time_series_group$avg_price.x +
                            time_series_group$total_volume.y* time_series_group$avg_price.y)/
               (time_series_group$total_volume.x+time_series_group$total_volume.y)) %>% 
  add_column(total_volume = (time_series_group$total_volume.x+time_series_group$total_volume.y))
time_series_group <- time_series_group %>% group_by(date) %>%  summarise(price = mean(adj_price), volume = mean(total_volume))

ts_reg <- lm(log(volume)~log(price), data = time_series_group)
summary(ts_reg)

ggplot(time_series_group, aes(x=log(volume)))+
  geom_point(aes(y = log(price)))
 
```
The time series log yields a coefficient estimate of -0.61305 with a standard 
error of .08337. This shows that there is a negative correlation between price 
and and volume. As the volume decreases, the price goes up and vice versa for 
increases in volume. For economic interpretation of this correlation, when the 
market is flooded with avocados, the price is lower because supply is high, and a 
shift in the supply and demand equilibrium occurs. Time series variation is likely 
a better estimate of supply side. This is because quantity supply shifts as 
producers are able to grow a varying amount based on climate conditions and 
dictates how many avocados can be supplied. 

## cross sectional series 

```{r echo = FALSE}
cross_sectional <- merge(organic, conventional, by = c("date", "region"))%>% 
  filter(region != "TotalUS" )
cross_sectional<- cross_sectional %>% 
  add_column(adj_price = (cross_sectional$total_volume.x* cross_sectional$avg_price.x +
                            cross_sectional$total_volume.y* cross_sectional$avg_price.y)/
               (cross_sectional$total_volume.x+cross_sectional$total_volume.y)) %>% 
  add_column(total_volume = (cross_sectional$total_volume.x+cross_sectional$total_volume.y))


cross_sectional_group <- cross_sectional %>% group_by(region) %>%  summarise(price = mean(adj_price), volume = mean(total_volume))

CS_reg <- lm(log(volume)~log(price), data = cross_sectional_group)
summary(CS_reg)

ggplot(cross_sectional_group, aes(x=log(volume)))+
  geom_point(aes(y = log(price)))

```
The cross sectional regression yields a coefficient of  -2.2287 with a standard
error of 1.0887. This shows that the price and quantity differs much more than 
time series as the coefficient is larger than that of the time series regression. 
When understanding the meaning of this coefficient, the more avocados sold, the
cheaper the price would be and across different regions, this change was quite
significant. The economic parameters that people would care about is that 
avocados distributed to different areas will affect the sale price and the higher 
the quantity delivered to those areas, on average, will reduce the sale price 
of the avocados. This is more driven by demand side factors as transportation 
around the US is relatively inexpensive so each area can, given a sufficient 
supply, provide enough avocados to meet the demand of the given area. 




## panel data regressions
```{r echo=FALSE}
panel <- merge(organic, conventional, by = c("date", "region"))%>% 
  filter(region != "TotalUS" )

panel<- panel %>% 
  add_column(adj_price = (panel$total_volume.x* panel$avg_price.x +
                           panel$total_volume.y* panel$avg_price.y)/
               (panel$total_volume.x+panel$total_volume.y)) %>% 
  add_column(total_volume = (panel$total_volume.x+panel$total_volume.y))

panel_reg <- lm(log(total_volume)~log(adj_price), data = panel)
summary(panel_reg)

```
In the panel regressions, the regression yielded a coefficient estimate of 
-1.34325 with a standard error of 0.05436. This fell between the coefficients 
of time series regression and cross section regression. This shows that an 
adjusted average of how the avocado market prices adjust with the change in 
avocado quantity. The economic parameters that people should care about from this
regression is that the price in avocados does change based on the quantity 
supplied for both supply and demand side reasons. Since this coefficient is 
closer to that of the cross series, demand side may play more of a factor on the
change in price and quantity of the avocados. 



### Part 4 - Elasticity of Supply

```{r echo = FALSE}
supply_elasticity <-time_series_group %>% 
  arrange(date) %>% 
  mutate(
    avg_price_prevdate = lag(price),
    price_delta = price - avg_price_prevdate,
    abs_value_price_delta = abs(price_delta),
    perc_price_change = (price -avg_price_prevdate)/avg_price_prevdate *100
    ) %>% 
  mutate(
    avg_vol_prevdate = lag(volume),
    vol_delta = volume - avg_vol_prevdate,
    abs_value_vol_delta = abs(vol_delta), 
    perc_vol_change = (volume - avg_vol_prevdate)/avg_vol_prevdate *100
    )
  #arrange(desc(abs_value_price_delta))

supply_elasticity_reg <- lm(perc_price_change~perc_vol_change,data = supply_elasticity)
summary(supply_elasticity_reg)

ggplot(supply_elasticity_reg, aes(x=perc_price_change))+
  geom_point(aes(y = perc_vol_change))+
  ggtitle("Supply Elasticity of Avocados - % change of quantity and price")






```

For understanding the elasticity of supply, I have applied the data set 
constructed for the time series regression as it covers how the supply of 
avocados change over time. Columns were added to find the % quantity change from 
previous date and % price change from previous dates to show how the change in
supply would affect the quantity and supply of avocados. The coefficient produced
by this regression is -0.37159. This means that there is some change in the 
quantity supplied based on price changes. The standard error band that I would 
give would be more than the regressions of 0.01736. The justification for the 
error band being greater than the regressions is that more factors affect the 
supply of avocados than just yearly climate trends. The graph shows the negative 
correlation between % change in volume and the % change in price, demonstrating 
to the department of agriculture how changes in supply affect these factors. 



### Part 5 - Elasticity of Demand


```{r echo = FALSE}
temperatureColor <- "#69b3a2"
priceColor <- rgb(0.2, 0.6, 0.9, 1)

demand_elasticity <-time_series_group %>% 
  mutate(
    year = year(date),
    month = month(date),
    day = day(date))%>%  filter(year == 2017, month >= 1, month < 3) %>% arrange(date) %>% mutate(
    avg_price_prevdate = lag(price),
    price_delta = price - avg_price_prevdate,
    abs_value_price_delta = abs(price_delta),
    perc_price_change = (price -avg_price_prevdate)/avg_price_prevdate *100
    ) %>% 
  mutate(
    avg_vol_prevdate = lag(volume),
    vol_delta = volume - avg_vol_prevdate,
    abs_value_vol_delta = abs(vol_delta), 
    perc_vol_change = (volume - avg_vol_prevdate)/avg_vol_prevdate *100
    )


ggplot(demand_elasticity, aes(x=perc_price_change))+
  geom_point(aes(y = perc_vol_change))



p1 <- demand_elasticity %>% 
  ggplot(aes(date, volume)) +
  geom_line(size=2, color=priceColor) 
p2 <- demand_elasticity %>% 
  ggplot(aes(date, price)) +
  geom_line(size=2, color=temperatureColor) 

demand_elasticity_reg <- lm(perc_price_change~perc_vol_change,data = demand_elasticity)
summary(demand_elasticity_reg)


p1  +
  ggtitle("# of avocados sold in Jan/Feb")+ 
  p2 +
  ggtitle("Price of avocados in Jan/Feb")


```
Given that the data supplied, there were no columns for avocados supplied and/or
avocados not sold that could contribute to calculating the elasticity of demand. 
Because of this, a natural experiment in the data could lead to an example of 
elasticity of demand. Looking at the trends in the data, there consistently is a
spike of avocados purchased in early February, which coincides with the 
super bowl. Google reveals that avocado demand spikes during this time because 
of this event. As such, this is a good natural experiment to determine the 
elasticity of demand. The regression yielded a coefficient estimate of -0.39907
and a standard error of 0.06323. Again, I believe that this standard error should
be greater than the given one because there are many factors that could impact
the demand side, in addition to the early Feb spike. The graphic produced for the
department of agriculture shows that, in early february, there is a giant spike 
for folume and simultaneously, there is a drop in price. 

### Part 6 - in line with elasticity of demand and supply


### Part 7 - Recommendations for further data collection

Overall, the data collected greatly presents it self as demand side data. It only
contains values for quantity sold which, theoretically, is at the equilibrium of 
supply and demand. More data collection could help assist understanding the 
efficiency of the market by providing data on how many avocados are supplied. 
This could take the form of farmers keeping track of how much they produce or 
it could also be tracked by how many avocados are not sold on the sellers/grocery
store's side of the supply chain. This would be beneficial as the time series
isn't a prefect representation of the supply. Now days, global transportation is 
cheap and freezing produce to be used in the off-season is common practice. As 
such, time series fails to recognize these effects in the supply of avocados. 







